---
title: "Statistical Learning Survey"
author: "Abraham Neuwirth"
date: "May 29, 2017"
output: 
  bookdown::pdf_document2:
     includes:
      in_header: preamble.tex
bibliography: bibliography.bib

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Introduction

Christopher Bishop [@pattern_recognition] in his seminal book "Pattern Recognition and Machine Learning", introduces the field of statistical learning with a classical story:

> [T]he extensive astronomical observations of Tycho Brahe in the 16th century allowed Johannes Kepler to discover the empirical laws of planetary motion, which in turn provided a springboard for the development of classical
mechanics.

It's the archtype of statistical learning success stories. Lot's of data, brilliant minds, and a model to illuminate and explain it all.

# Data Representation

## Vectors

## Matrices

## Tensors

# Supervised Learning

```{lemma}
This is a lemma
```

## Least-Squares Regression

## Support Vector Machines

# Unsupervised Learning

## Classification

## Naive Bayes

## K-Nearest Neighbors

# Regularlization
Overfitting is a problem. regularlization penalizes biggest predictors. "Regularization can be accomplished by restricting the hypothesis space $\mathcal {H}$" 

## Tikhonov regularization (Ridge Regression)


## Lasso Regression


## Principal Components
Dimensionality reduciton

# Reinforcment Learning

# Deep Learning

This is a level 1 heading



## Level 2

Blah blah [se also @islr ch. 1] and @eslr.

### equations
This s an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>. See equation \@ref(eq:sum) there's also \@ref(eq:sum2)

\begin{equation} 
  \sum_{i}^{x}5+1
  (\#eq:sum)
\end{equation} 


\begin{equation} 
  \sum_{i}^{x}5+5
  (\#eq:sum2)
\end{equation} 

### Level 4

```{theorem}
Here is A theorem.
```


## Including Plots

You can also embed plots, for example:

```{r pressure,  echo=FALSE, fig.cap="test a caption"}
library(ggplot2)
ggplot(pressure, aes(temperature, pressure)) + geom_point()
```

Note that the `echo = FALSE` parameter was added to the code chunk \@ref(fig:pressure) to prevent printing of the R code that generated the plot. @pattern_recognition

also, lets put this in here

\begin{equation} 
\begin{split}
\mathrm{Var}(\hat{\beta}) & =\mathrm{Var}((X'X)^{-1}X'y)\\
 & =(X'X)^{-1}X'\mathrm{Var}(y)((X'X)^{-1}X')'\\
 & =(X'X)^{-1}X'\mathrm{Var}(y)X(X'X)^{-1}\\
 & =(X'X)^{-1}X'\sigma^{2}IX(X'X)^{-1}\\
 & =(X'X)^{-1}\sigma^{2}
\end{split}
(\#eq:var-beta)
\end{equation} 


and now lets referene \@ref(eq:var-beta) for god sake
Note that the `echo = FALSE` parameter was added to the code chunk \@ref(fig:pressure) to prevent printing of the R code that generated the plot.

Note that the `echo = FALSE` parameter was added to the code chunk \@ref(fig:pressure) to prevent printing of the R code that generated the plot.

Note that the `echo = FALSE` parameter was added to the code chunk \@ref(fig:pressure) to prevent printing of the R code that generated the plot.
Note that the `echo = FALSE` parameter was added to the code chunk \@ref(fig:pressure) to prevent printing of the R code that generated the plot.

## level  chapeter

fdsfdsf sdf
fsd f
sdf 
sd
f 
sf

# let's try  an ew cahpeter 

\begin{equation} 
\begin{split}
\mathrm{Var}(\hat{\beta}) & =\mathrm{Var}((X'X)^{-1}X'y)\\
 & =(X'X)^{-1}X'\mathrm{Var}(y)((X'X)^{-1}X')'\\
 & =(X'X)^{-1}X'\mathrm{Var}(y)X(X'X)^{-1}\\
 & =(X'X)^{-1}X'\sigma^{2}IX(X'X)^{-1}\\
 & =(X'X)^{-1}\sigma^{2}
\end{split}
(\#eq:var-gamma)
\end{equation} 
trterter \@ref(eq:var-gamma)

# References
